---
title: "covid 19"
author: "Giulianni P"
date: "Fall 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex


Based on the gender we'll be using the CAT~CAT helper file.

Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```
Out of the 1800 Females in Korea during Covid 19 there were far less deaths compared to that of males.


## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`
```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  
Based on the barplot shown above you're more likely to be a deceased male in Korea that has died due to covid 19.




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The probability of a Male dying from covid 19 compared to that of a female was by 2.6 times



## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P-value is .0012 which would be lower than .05 allowing us to determine that we reject the null hypothesis.

## Conclusions

Question 1 - Is dying independent of sex

Going back to answer the first question, we can say for this study alone it showed us that Sex was an impotand and determining factor.


